Trait-matching simulation

### 2016-12-01 SimulatedData.Rmd
trait-matching knitted

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(pez)
library(ggplot2)
library(knitr)
library(dplyr)
library(stringr)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
```

@@

size_x = Number of level 1 (eg. Plants)
size_y = Number of level 2 (eg. Pollinators)
traits_x = Trait frame for level 1
traits_y = Trait frame for level 2
phylogeny_x = phylogeny for level 1
phylogeny_y = phylogeny for level 2
Obs = a x by y matrix of visits

```{r}
#sim_assemblages<-function(x=,y=,traits=
## Process Model
```

```{r,fig.height=5,fig.width=8,eval=T}
#Number of consumer species
h_species=10
plant_species=10
cameras<-5
days<-3

#Bill sizes
Bill<-rpois(h_species,10)

#Corolla sizes
Corolla<-rpois(plant_species,15)

#Subtract both and take absolute value, convert cm
traitmatch<-abs(sapply(Corolla,function(x) x - Bill)/10)

@@

#abundance

#species variance in slopes
beta1_sigma<- 0.2

#Species alpha_mu 
alpha_mu<- 3
alpha_sigma<- 0.2

#Poisson overdipersion
tauE<-10

#species level
detection= inv.logit(rnorm(h_species,0,1.67))
beta1<-rnorm(h_species,beta1_mu,beta1_sigma)
alpha<-rnorm(h_species,alpha_mu,alpha_sigma)
```

#Compute true interaction matrices

```{r,eval=T}
#for each species loop through and create a replicate dataframe
obs<-array(dim=c(h_species,plant_species,cameras,days))
lambda<-array(dim=c(h_species,plant_species,cameras))
N<-array(dim=c(h_species,plant_species,cameras))

#create intensities
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
      lambda[x,y,z]<-exp(alpha[x] + beta1[x] * traitmatch[x,y] + rnorm(1,0,1/sqrt(tauE)))
  }
  }
}

#draw latent states
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
      # true latent count
      N[x,y,z]<-rpois(1,lambda[x,y,z])
    }
  }
}
```

##View correlation in simulated latent state

```{r}
mdat<-melt(N)
colnames(mdat)<-c("Bird","Plant","Camera","Interactions")

traitmelt<-melt(traitmatch)
colnames(traitmelt)<-c("Bird","Plant","traitmatch")

mdat<-merge(mdat,traitmelt,c("Bird","Plant"))
ggplot(mdat,aes(x=traitmatch,y=Interactions,col=as.factor(Bird))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args = list(family = "poisson")) + labs(col="Bird") + xlab("Absolute value of Bill Length - Corolla Length ")
```

##View Detection Rates

```{r}
obs.state<-melt(obs)
colnames(obs.state)<-c("Bird","Plant","Camera","Day","Yobs")
obs.state<-merge(mdat,obs.state,by=c("Bird","Plant","Camera"))
ggplot(obs.state,aes(x=Interactions,y=Yobs,col=Camera)) + geom_point() + theme_bw() + geom_abline() + coord_equal()
```

>>>>

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(bipartite)
library(ggplot2)
library(knitr)
library(dplyr)
library(stringr)
library(R2jags)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=5,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
```

@@

size_x = Number of level 1 
size_y = Number of level 2
traits_x = Trait frame for level 1
traits_y = Trait frame for level 2
Obs = x by y matrix of visits

```{r}
simA<-function(size_x,size_y,traits_x,traits_y,beta1_sigma= 0.4,alpha_sigma= 0){

  #Subtract both and take absolute value, convert cm
  traitmatch<-abs(sapply(traits_y,function(x) x - traits_x))

@@

  #species variance in slopes
  beta1_sigma<- 0.1

  #Species alpha_mu 
  alpha_mu<- 3
  alpha_sigma<- 0

  #species level
  beta1<-rnorm(size_x,beta1_mu,beta1_sigma)
  alpha<-rnorm(size_x,alpha_mu,alpha_sigma)

  #for each species loop through and create a replicate dataframe
  obs<-array(dim=c(size_x,size_y))
  N<-array(dim=c(size_x,size_y))

  #create intensities
  for(x in 1:size_x){
    for (y in 1:size_y){

        #intensity
        N[x,y]<-exp(alpha[x] + beta1[x] * traitmatch[x,y])

        #draw one state
        obs[x,y]<-rpois(1,N[x,y])
    }
  }

#draw intensity

tx<-data.frame(I=1:length(traits_x),TraitI=traits_x)
ty<-data.frame(J=1:length(traits_y),TraitJ=traits_y)

#view trait matching
dat<-melt(obs)
colnames(dat)<-c("I","J","Interactions")

dat<-merge(dat,tx)
dat<-merge(dat,ty)

#define data types, make species
dat$I<-letters[dat$I]
dat$J<-letters[dat$J]

return(dat)
}
```

```{r}
size_x = 5
size_y = 19

#Bill sizes
traits_x<-rpois(size_x,10)

#Corolla sizes
traits_y<-rpois(size_y,10)

dat<-simA(size_x=size_x,size_y=size_y,traits_x=traits_x,traits_y=traits_y)

#view heatmap
heatmap<-function(dat){
  ggplot(data=dat,aes(x=I,y=J,fill=Interactions)) + geom_tile() +    scale_fill_continuous(low="white",high="red")
}

heatmap(dat)

traitplot<-function(dat){
    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="poisson"))
}

traitplot(dat)
```

# Poisson Regression

```{r,eval=T}
#format traitmatch as matrix
dat$Traitmatch<-abs(dat$TraitI-dat$TraitJ)
Traitmatch<-acast(data=dat,I~J,value.var="Traitmatch")

runs<-100000

#Source model
source("Bayesian/Poisson.R")

#print model
print.noquote(readLines("Bayesian//Poisson.R"))

  #for parallel run
  Yobs=dat$Interactions
  Bird=as.numeric(as.factor(dat$I))
  Plant=as.numeric(as.factor(dat$J))
  Birds=length(unique(dat$I))
  Traitmatch=Traitmatch
  Plants=length(unique(dat$J))
  Nobs<-length(Yobs)

  #Parameters to track
  ParsStage <- c("beta1","ynew","fit","fitnew")

  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 1  #thinning rate
  nb <- max(0,runs-500) # number to discard for burn-in
  nc <- 2  # number of chains

  modelDat<-list("Yobs","Bird","Plant","Plants","Traitmatch","Birds","Nobs")

    m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Poisson.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
```

```{r}
parsO<-melt(m1$BUGSoutput$sims.array)
colnames(parsO)<-c("Draw","Chain","parameter","estimate")

#label species and plants
l<-levels(parsO$parameter)

#parameters to save
totrack<-m1$parameters.to.save

#assign species index to ragged frame.
sp_pl<-data.frame(parameter=l,Index=as.numeric(str_match(l,pattern="\\[(\\d+)]")[,2]),par=str_extract(l,"\\w+"))

#merge levels
pars<-merge(parsO,sp_pl)

pars<-pars[!pars$par %in% "deviance",]

```

```{r,fig.width=15}
#view chaings
ggplot(pars[pars$par %in% c("beta1"),],aes(x=Draw,col=as.factor(Chain),y=estimate))  + geom_line() + facet_wrap(~Index,scales="free") 
```

##Posterior predictive check

```{r,fig.height=4,fig.width=8}
fitstat<-droplevels(pars[pars$par %in% c("fit","fitnew"),])
fitstat<-dcast(fitstat,Draw+Chain~par,value.var="estimate")

#add 1:1 line
ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point() + theme_bw() + coord_equal()
psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
psim4

```

#View simulated and predicted traitmatching

```{r}

#merge the dataframes.
dat$Index<-1:nrow(dat)

#posterior distribution for each index
ypred<-pars %>% filter(par=="ynew") %>% group_by(Index) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),higher=quantile(estimate,0.95))

datpred<-merge(dat,ypred)

ggplot(datpred) + geom_point(aes(x=Traitmatch,y=Interactions),col="black") + geom_ribbon(aes(x=Traitmatch,ymin=higher,ymax=lower),alpha=0.6) + theme_bw()

ggplot(datpred) + geom_point(aes(x=Traitmatch,y=Interactions),col="black") + geom_ribbon(aes(x=Traitmatch,ymin=higher,ymax=lower),alpha=0.6) + theme_bw() + facet_wrap(~I)
```

# Compute emergent statistics

```{r}

makeN<-function(x){
  predweb<-acast(data=x,I~J,value.var="estimate")

  #calculate network statistic
  nstat<-networklevel(predweb,index=c("weighted connectance","weighted NODF"))
  data.frame(Metric=names(nstat),nstat)
}

#calculate network stats for each model
nstats<-pars %>% filter(par=="ynew") %>% group_by(Draw,Chain) %>% inner_join(dat) %>% do(makeN(.))

#calculate true network stat
tweb<-acast(dat,I~J,value.var="Interactions")
tstat<-networklevel(tweb,index=c("weighted connectance","weighted NODF"))
tstat<-data.frame(Metric=names(tstat),tstat)

ggplot(nstats,aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=2) + theme_bw()  + geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) + labs(x="value")
```
